---
title: "p8105_hw1_mw3845"
author: "Minghe Wang"
date: "2024-09-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

```{r problem1 - load dataset penguins}
data("penguins", package = "palmerpenguins")
```

The dataset `penguins` contains 8 variables including `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, `year`. This dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean of `flipper_length_mm` is `r mean(penguins$flipper_length_mm, na.rm=TRUE)`.

```{r problem1 - scatterplot_flipperLength_vs_billLength}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point() +
  labs(title = "Bill length(mm) vs Flipper length(mm)")

ggsave('bill vs flipper.png')
```
## Problem 2

###### Generating dataframe containing different types of variables:
```{r problem2 - generate 4x10 dataframe}
library(tidyverse)
set.seed(123)

#generate 4 types of variables
standard_normal_vec = rnorm(10, mean = 0, sd = 1)
positive_logical_vec = standard_normal_vec > 0
character_vec = c('milk', 'cheese', 'egg', 'banana', 'apple', 'toast', 'orange juice', 'peanut butter', 'cucumber',  'coffee')
factor_vec = factor(sample(c('low', 'medium', 'high'), size = 10, replace = TRUE))

#merge variables into the dataframe
df = data.frame(
  standard_normal_var = standard_normal_vec,
  positive_logical_var = positive_logical_vec,
  character_var = character_vec,
  factor_var = factor_vec
)
```

###### Mean of each variable in dataframe

For thses four types of variables, we can only take mean of `standard_normal_var` and `positive_logical_var` because `character_var` and `factor_var` are not numerical nor logical.

```{r problem2 - taking mean of each variable}
#taking mean of 4 varibales
df %>% pull(standard_normal_var) %>% mean()
df %>% pull(positive_logical_var) %>% mean()
df %>% pull(character_var) %>% mean() #FAILED
df %>% pull(factor_var) %>% mean() #FAILED
```

###### Convert variables to numeric:

Here we attempt to convert `positive_logical_var`, `character_var`, and `factor_var` to numeric variable using `as.numeric`:

Logical and factor variable can be converted to numerical variables, while the character variable are converted into NAs.

- `positive_logical_var` can be converted to numeric because R treats logical values as `TRUE=1` and `FALSE=0`.
- `character_var` cannot be converted into numeric because characters cannot be coerced into numeric values without some meaningful mapping for R to interpret.
- `factor_var` can be converted to numeric because R will store factor values as integer internally, and `as.numeric` can convert the factor values into their corresponding integers.

Above also explain why we will get NA while taking mean of character and factor variable. For factor variable, simply taking mean of it will result in NA because factor variables are not inherently available for numeric operations although R will assign integer to each level of the factor variables.

```{r problem2 - convert variables to numeric, echo = TRUE, eval = FALSE}
#convert logical, character, and factor variables into numeric type
df %>% pull(positive_logical_var) %>% as.numeric()
df %>% pull(character_var) %>% as.numeric()
df %>% pull(factor_var) %>% as.numeric()
```